---
title: "Get started with openetR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with openetR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(openetR)
```

`openetR` provides a small, consistent interface to OpenET's raster API. It
uses [httr2](https://httr2.r-lib.org/) for HTTP requests and returns decoded
JSON responses without requiring a web browser or an Earth Engine client.

## Authenticate

Create an API key in the [OpenET account dashboard](https://account.etdata.org/
settings/api), then set it as an environment variable. This keeps credentials
out of scripts, Git history, and rendered reports.

```{r, eval = FALSE}
Sys.setenv(OPENET_API_KEY = "your-api-key")
```

The request functions read this variable automatically. For temporary or
programmatic credentials, pass `key` directly instead.

## Retrieve a point time series

OpenET uses WGS84 longitude and latitude coordinates. The following request
retrieves monthly ensemble ET at a point in California. API requests are not
run while building this vignette, so the example remains reproducible without
credentials.

```{r, eval = FALSE}
et <- openet_point_timeseries(
  longitude = -121.36322,
  latitude = 38.87626,
  start = "2020-01-01",
  end = "2020-12-31",
  interval = "monthly",
  model = "ensemble",
  variable = "et",
  reference_et = "gridmet",
  units = "mm"
)
```

Use `interval = "daily"` for daily data. OpenET currently limits daily point
and polygon requests to ten years. Available model, variable, and reference ET
values are documented by [OpenET](https://openet.gitbook.io/docs/reference/api-reference).

## Inspect collection metadata

Use `openet_metadata()` to inspect the processing metadata for a location and
collection before interpreting a result.

```{r, eval = FALSE}
metadata <- openet_metadata(
  geometry = c(-121.36322, 38.87626),
  model = "ensemble",
  variable = "et"
)
```
