---
title: "Get started with openetR"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
vignette: >
  %\VignetteIndexEntry{Get started with openetR}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

`openetR` provides a compact interface to the [OpenET raster
API](https://openet.gitbook.io/docs/reference/api-reference/raster). Retrieve
evapotranspiration (ET) time series for a point or polygon, then use metadata
to understand the selected raster collection.

```{r}
#| label: load-openetr
library(openetR)
```

## Authenticate {#sec-authenticate}

Create an API key in the [OpenET account dashboard](https://account.etdata.org/settings/api).
Store it in an environment variable so it stays out of scripts, Git history,
and rendered reports.

```{r}
#| eval: false
Sys.setenv(OPENET_API_KEY = "your-api-key")
```

All request functions read this variable automatically. For temporary or
programmatic credentials, supply `key` directly instead.

::: {.callout-important}
## Keep keys private

Do not commit an API key to a repository. Prefer environment variables or a
secret-management system for deployed work.
:::

## Retrieve a point time series {#sec-point}

OpenET uses WGS84 longitude and latitude coordinates. The following request
retrieves monthly ensemble ET at a point in California. It is shown but not run
so this article can build without credentials.

```{r}
#| eval: false
et <- openet_point_timeseries(
  longitude = -121.36322,
  latitude = 38.87626,
  start = "2020-01-01",
  end = "2020-12-31",
  interval = "monthly",
  model = "ensemble",
  variable = "et",
  reference_et = "gridmet",
  units = "mm"
)
```

Use `interval = "daily"` for daily values. Daily point and polygon requests
are limited to ten years by OpenET. See the API documentation for supported
models, variables, reference ET sources, and quotas.

## Inspect collection metadata {#sec-metadata}

Use `openet_metadata()` to inspect a collection at a location before
interpreting a result.

```{r}
#| eval: false
metadata <- openet_metadata(
  geometry = c(-121.36322, 38.87626),
  model = "ensemble",
  variable = "et"
)
```

For area summaries, continue to the [polygon workflow](polygon-workflows.html).
