---
title: "Work with OpenET polygons"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
vignette: >
  %\VignetteIndexEntry{Work with OpenET polygons}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

Use `openet_polygon_timeseries()` to summarize an OpenET raster collection for
a simple polygon. Coordinates must use WGS84 (EPSG:4326) and are supplied as
alternating longitude/latitude values.

```{r}
#| label: load-openetr
library(openetR)
```

## Define a polygon {#sec-geometry}

This rectangle does not repeat its first coordinate because OpenET accepts a
simple sequence of vertex pairs. Use a closed ring only if your source data
requires it; `openetR` passes coordinates to OpenET unchanged.

```{r}
field <- c(
  -121.00747, 44.24420,
  -121.00747, 44.24742,
  -121.00295, 44.24742,
  -121.00295, 44.24422
)
```

## Request an aggregated time series {#sec-request}

Choose the pixel aggregation with `reducer`. `"mean"` is appropriate for an
average depth across an area. Use `"sum"` only when its units and
interpretation are appropriate for the selected variable.

```{r}
#| eval: false
field_et <- openet_polygon_timeseries(
  geometry = field,
  start = "2021-01-01",
  end = "2021-12-31",
  interval = "monthly",
  model = "ensemble",
  variable = "et",
  reference_et = "gridmet",
  reducer = "mean",
  units = "mm"
)
```

## Prevent common input errors {#sec-validation}

`openetR` checks coordinate bounds, requires an even number of coordinate
values, and verifies that dates are ordered before making a request.

::: {.callout-tip}
## Start with metadata

Call `openet_metadata(field)` before a long analysis to confirm the selected
model, variable, reference ET source, and collection version.
:::

For larger collections of fields, use OpenET's multipolygon or export endpoints
directly until a high-level client is added to `openetR`.
