---
title: "Work with OpenET polygons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Work with OpenET polygons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(openetR)
```

Use `openet_polygon_timeseries()` to summarize an OpenET raster collection for
a simple polygon. Coordinates must be WGS84 and are supplied as alternating
longitude/latitude values.

## Define a polygon

This rectangle is shown without repeating its first coordinate because OpenET
accepts a simple sequence of vertex pairs. Use a closed ring when your source
data requires it; `openetR` passes the coordinates to OpenET unchanged.

```{r}
field <- c(
  -121.00747, 44.24420,
  -121.00747, 44.24742,
  -121.00295, 44.24742,
  -121.00295, 44.24422
)
```

## Request an aggregated time series

Choose the pixel aggregation with `reducer`. `"mean"` is appropriate for an
average depth across an area, while `"sum"` is useful for totals only when its
units and interpretation match the selected OpenET variable.

```{r, eval = FALSE}
field_et <- openet_polygon_timeseries(
  geometry = field,
  start = "2021-01-01",
  end = "2021-12-31",
  interval = "monthly",
  model = "ensemble",
  variable = "et",
  reference_et = "gridmet",
  reducer = "mean",
  units = "mm"
)
```

## Validate inputs before a request

`openetR` validates coordinate bounds, requires an even number of coordinate
values, and checks that the requested date range is ordered. Those checks catch
common errors locally, before spending an API request. For larger collections
of fields, use OpenET's multipolygon or export endpoints directly until a
high-level client is added to `openetR`.
